%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%                                                                 %
%                         ACKNOWLEDGEMENT                         %
%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 
\specialhead{ACKNOWLEDGMENT}
 
%This dissertation is the essence of the not-so-short four and half years' work in the Optimal Design Lab of Professor Jason E. Hicken at Rensselaer Polytechnic Institute.  The PhD experience has taught me numerous valuable lessons in doing research,  given me sharper insights in numerical optimization field, and confidence in doing whatever is waiting ahead.  

I would like to thank my advisor, Professor Jason E. Hicken, for the numerous discussions he patiently had with me 
for the past five years. His broad knowledge and insights in numerical simulation and design field continue to motivate me.  Beyond being professional, he is a very nice person to work with, and it is my pleasure to have him as my advisor. 

I thank my doctoral committee members, Prof. Assad A. Oberai, Prof. Lucy T. Zhang, Prof. John E. Mitchell for the time they spent on my candidacy and thesis defense. Their valuable inputs greatly help improve the quality of my thesis work. 

Many thanks to my fellow lab mates in CII lab for providing a lively and humorous environment: Alp Dener, Anthony Ashley,  Jared Crean, Jianfeng Yan, Kinshuk Panda. In particular, thank you to Alp Dener, whose work in Kona laid the foundation for my thesis research, and who very generously gave me numerous valuable suggestions. 

I gratefully acknowledge Dr. Gaetan Kenway, Dr. Charles (Sandy) Mader in MDO design lab at University of Michigan for their help with the CFD solver ADflow. 

Lastly I am grateful to have my family with me, Yongjian, and most recently Emma, my mom and my sister. 

%I am indebted to my mom for her constant support and love, despite the long distance and long time separation caused by my studying in the U.S.. My only sister, for the special bond we had through the rough days, and for your love and trust in me all the time. 
%
%Yongjian Yang, thank you for being my life companion, and the meanings and purposes you instill to my life. I am grateful for having my baby daughter Emma. Everyday is a surprise with you. The lively energy you emits is my greatest source of happiness and inspiration.  

This thesis is jointly funded by the NASA Learn project, Prof. Hicken's startup grant, and TA funding from the MANE department. Specially, thanks to OGE at RPI for supporting me with Childbirth Accommodation in fall 2017 after my baby was born, which enable me to spend more time with my baby and smoothly finish my PhD. 
